---
title: Identifying the best predictive biomarker in pharmacogenomics through multiple comparisons with the best
output:
  html_document: default
  mainfont: Times New Roman
  fontsize: 12
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## LOAD PACKAGES AND FUNCTIONS

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(data.table)
library(glmnet)
library(gridExtra)
```

```{r}
dir <- "~/PROJECT/Dissertation/Code/" # CHANGE IT TO YOUR OWN PATH

source(paste0(dir,"Functions/Functions-IT_p1.R"))
source(paste0(dir,"Functions/Functions-IT_p2.R"))
source(paste0(dir,"Functions/Functions-Ranking.R"))
source(paste0(dir,"Functions/Functions-LSmeans+MCB.R"))
source(paste0(dir,"Functions/Functions-Fig.R"))
```

## LOAD THE GENOTYPE DATA

```{r}
load(paste0(dir,"Data/geno.Rdata"))
```

## PARAMETERS SETUP

```{r}
gamma = 1
alpha = -6
BETA = seq(1, 3, 0.5)
N <- c(100, 379)
index.n100 <- c(1:97, which(g1 != g3))
NREP = 1:500
```

## SET SEED

```{r}
set.seed(1002)
```

## FIGURE 3A 

rs429358 is the causal SNP that determine the drug response. The power is defined as the probability that rs429358 is identified as the best predictive SNP.

```{r}
Tr <- rbinom(nrow(G), 1, 0.5)

Gx <- cbind.data.frame("chr19.rs429358.pos45411941"=g1, "chr19.rs7412.pos45412079"=g2, G[,2301:2303]) %>% as.matrix()

sigma = 3
```

```{r}
p11 <- p12 <- p13 <- p14 <- p21 <- p22 <- p23 <- p24 <- matrix(NA, ncol = length(BETA), nrow = length(NREP))

for (n in N) {
  for (k in 1:length(BETA)) {
    beta = BETA[k]
    print(paste0("n = ",n,", beta = ",beta))
    for (i in 1:length(NREP)) {
      Y <- gamma*Tr + alpha*g1 + beta*(Tr*g1) + rnorm(nrow(Gx), 0, sigma)
      
      if(n == 100){ 
        Trx <- Tr[index.n100]; Gxx <- Gx[index.n100,]; Yx <- Y[index.n100]
      
        re1 <- LSmeans_MCB_est(Yx, Trx, Gxx)
        re2 <- pvalue.rank(Yx, Trx, Gxx)
        re3 <- Lasso(Yx, Trx, Gxx)
        re4 <- IT(Yx, Trx, Gxx)
        
        p11[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941")) 
        p12[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941")) 
        p13[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941")) 
        p14[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941")) 
      }
    
      if(n == 379){ 
        re1 <- LSmeans_MCB_est(Y, Tr, Gx)
        re2 <- pvalue.rank(Y, Tr, Gx)
        re3 <- Lasso(Y, Tr, Gx)
        re4 <- IT(Y, Tr, Gx)
      
        p21[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941"))
        p22[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941"))
        p23[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941"))
        p24[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941"))
      }
    }
  }
}

colnames(p11) <- colnames(p12) <- colnames(p13) <- colnames(p14) <- colnames(p21) <- colnames(p22) <- colnames(p23) <- colnames(p24) <- BETA

power <- cbind(p11,p12,p13,p14,p21,p22,p23,p24)
```

```{r}
RE1 <- cbind.data.frame(n = rep(N, each=4*length(BETA)),
                        beta = rep(BETA, 4*length(N)),
                        method = rep(rep(c("3M","P-value Ranking","Lasso","IT"),each=length(BETA)),length(N)),
                        mean = apply(power, 2, mean))
```

## FIGURE 3B

rs429358 is the causal SNP that determine the drug response. rs4420638 (which is in high LD to rs429358) is also among the candidates. The power is defined as the probability that rs429358 or rs4420638 is identified as the best predictive SNP.

```{r}
Gx <- cbind.data.frame("chr19.rs429358.pos45411941"=g1, "chr19.rs7412.pos45412079"=g2, "chr19.rs4420638.pos44564995"=g3, G[,2301:2302]) %>% as.matrix()

sigma = 3
```

```{r}
p11 <- p12 <- p13 <- p14 <- p21 <- p22 <- p23 <- p24 <- matrix(NA, ncol = length(BETA), nrow = length(NREP))

for (n in N) {
  for (k in 1:length(BETA)) {
    beta = BETA[k]
    print(paste0("n = ",n,", beta = ",beta))
    for (i in 1:length(NREP)) {
      Y <- gamma*Tr + alpha*g1 + beta*(Tr*g1) + rnorm(nrow(Gx), 0, sigma)
      
      if(n == 100){ 
        Trx <- Tr[index.n100]; Gxx <- Gx[index.n100,]; Yx <- Y[index.n100]
        
        re1 <- LSmeans_MCB_est(Yx, Trx, Gxx)
        re2 <- pvalue.rank(Yx, Trx, Gxx)
        re3 <- Lasso(Yx, Trx, Gxx)
        re4 <- IT(Yx, Trx, Gxx)
        
        p11[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995")) 
        p12[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995")) 
        p13[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
        p14[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
      }
    
      if(n == 379){ 
        re1 <- LSmeans_MCB_est(Y, Tr, Gx)
        re2 <- pvalue.rank(Y, Tr, Gx)
        re3 <- Lasso(Y, Tr, Gx)
        re4 <- IT(Y, Tr, Gx)
        
        p21[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
        p22[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
        p23[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
        p24[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
      }
    }
  }
}

colnames(p11) <- colnames(p12) <- colnames(p13) <- colnames(p14) <- colnames(p21) <- colnames(p22) <- colnames(p23) <- colnames(p24) <- BETA

power <- cbind(p11,p12,p13,p14,p21,p22,p23,p24)
```

```{r}
RE2 <- cbind.data.frame(n = rep(N, each=4*length(BETA)),
                        beta = rep(BETA, 4*length(N)),
                        method = rep(rep(c("3M","P-value Ranking","Lasso","IT"),each=length(BETA)),length(N)),
                        mean = apply(power, 2, mean))
```

## FIGURE 3C

The power is defined as the probability that rs429358 is identified as the best predictive SNP.

```{r}
Gx <- cbind.data.frame("chr19.rs429358.pos45411941"=g1, "chr19.rs7412.pos45412079"=g2, G[,2301:2303]) %>% as.matrix()

sigma = 6
```

```{r}
p11 <- p12 <- p13 <- p14 <- matrix(NA, ncol = length(BETA), nrow = length(NREP))

for (k in 1:length(BETA)) {
  beta = BETA[k]
  print(paste0("beta = ",beta))
  for (i in 1:length(NREP)) {
    Y <- gamma*Tr + alpha*g1 + beta*(Tr*g1) + rnorm(nrow(Gx), 0, sigma)
    
    re1 <- LSmeans_MCB_est(Y, Tr, Gx)
    re2 <- pvalue.rank(Y, Tr, Gx)
    re3 <- Lasso(Y, Tr, Gx)
    re4 <- IT(Y, Tr, Gx)
    
    p11[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941"))
    p12[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941"))
    p13[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941"))
    p14[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941"))
  }
}

colnames(p11) <- colnames(p12) <- colnames(p13) <- colnames(p14) <- BETA

power <- cbind(p11,p12,p13,p14)
```

```{r}
RE3 <- cbind.data.frame(n = rep(379, 4*length(BETA)),
                        beta = rep(BETA, 4),
                        method = rep(c("3M","P-value Ranking","Lasso","IT"),each=length(BETA)),
                        mean = apply(power, 2, mean))
```

## FIGURE 3D 

The power is defined as the probability that rs429358 or rs4420638 (in high LD to rs429358) is identified as the best predictive SNP.

```{r}
Gx <- cbind.data.frame("chr19.rs429358.pos45411941"=g1, "chr19.rs7412.pos45412079"=g2, "chr19.rs4420638.pos44564995"=g3, G[,2301:2302]) %>% as.matrix()

sigma = 6
```

```{r}
p11 <- p12 <- p13 <- p14 <- matrix(NA, ncol = length(BETA), nrow = length(NREP))

for (k in 1:length(BETA)) {
  beta = BETA[k]
  print(paste0("beta = ",beta))
  for (i in 1:length(NREP)) {
    Y <- gamma*Tr + alpha*g1 + beta*(Tr*g1) + rnorm(nrow(Gx), 0, sigma)
    
    re1 <- LSmeans_MCB_est(Y, Tr, Gx)
    re2 <- pvalue.rank(Y, Tr, Gx)
    re3 <- Lasso(Y, Tr, Gx)
    re4 <- IT(Y, Tr, Gx)
    
    p11[i,k] <- as.numeric(names(re1)[which.max(re1)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
    p12[i,k] <- as.numeric(names(re2)[which.max(re2)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
    p13[i,k] <- as.numeric(names(re3)[which.max(re3)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
    p14[i,k] <- as.numeric(names(re4)[which.max(re4)] %in% c("chr19.rs429358.pos45411941","chr19.rs4420638.pos44564995"))
  }
}

colnames(p11) <- colnames(p12) <- colnames(p13) <- colnames(p14) <- BETA

power <- cbind(p11,p12,p13,p14)
```

```{r}
RE4 <- cbind.data.frame(n = rep(379, 4*length(BETA)),
                        beta = rep(BETA, 4),
                        method = rep(c("3M","P-value Ranking","Lasso","IT"),each=length(BETA)),
                        mean = apply(power, 2, mean))
```

## FIGURE 3 (COMBINED)

```{r warning=FALSE}
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

px <- power_comp_legend(RE1, title="A", legend.position="bottom")

mylegend <- g_legend(px)
```

```{r warning=FALSE}
p1 <- power_comp(RE1, title="A", legend.position="none")
p2 <- power_comp(RE2, title="B", legend.position="none")
p3 <- power_comp(RE3, title="C", legend.position="none")
p4 <- power_comp(RE4, title="D", legend.position="none")

p0 = grid.arrange(p1, p2, p3, p4, mylegend, layout_matrix = rbind(c(1,1,3), c(2,2,4), c(5,5,5)), heights = c(5,5,1))
```

```{r}
dir2 <- "~/PROJECT/Dissertation/Code/Results/" # CHANGE IT TO YOUR OWN PATH

ggsave(filename = paste0(dir2,"Fig3.png"), plot = p0,
       device = "png",
       width = 40,
       height = 35,
       units = "cm",
       dpi = 300,
       limitsize = FALSE)
```

